---
title: "Prediction Assignment Writeup"
author: "Alexander Oleynik"
date: "27 February 2016"
output: html_document
---
```{r setup, echo=FALSE, warning=FALSE, message = FALSE}
library(knitr)
options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, cache=TRUE)
library(caret)
```
#Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Data

Download dataset
```{r cache=TRUE}
setwd("C:\\Users\\Alex\\YandexDisk\\Documents\\Coursera\\Machine Learning\\Week4")
if (!file.exists("pml-training.csv")) {
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
                  destfile = "pml-training.csv")
}
if (!file.exists("pml-testing.csv")) {
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
                  destfile = "pml-testing.csv")
}
```

#Preprocessing

Some values of the objects in the dataset are NA. They need to be excluded from dataset.
```{r}
pml_training<-read.csv("pml-training.csv", header=T, na.strings=c("NA", "#DIV/0!"))
pml_testing<-read.csv("pml-testing.csv", header=T, na.string=c("NA", "#DIV/0!"))

namesData <- names(pml_training)

pml_trainingTidy <- pml_training
pml_testingTidy <- pml_testing

for (name in namesData) {
  num_na <- sum(is.na(pml_training[, name]))
  if (num_na > 0){
    pml_trainingTidy[,name] <- NULL
    pml_testingTidy[,name] <- NULL
  }
}
```

The values that are not required for the prediction (username, timestamp, ...) should also be deleted.
```{r}
pml_trainingTidy <- pml_trainingTidy[-(1:7)]
pml_testingTidy <- pml_testingTidy[-(1:7)]
```

#Modelling

Divide data into train and test.
```{r cache=TRUE, warning=FALSE, message = FALSE}
library(caret)

inTrain<-createDataPartition(y=pml_trainingTidy$classe, p=0.6,list=F)
training<-pml_trainingTidy[inTrain,] 
test<-pml_trainingTidy[-inTrain,] 
```

Use random forest classifier as algorithm of clasification. To increase productivity of modelling use doParallel library. 
To improve quality make preprocessing and use cross validation with number=5. 
```{r cache=TRUE}
 library(doParallel)
 cl <- makeCluster(detectCores())
 registerDoParallel(cl)

set.seed(12345)
fitControl<-trainControl(method="cv", number=5, allowParallel=T, verbose=T)
rffit<-train(classe~.,data=training, method="rf", trControl=fitControl, 
             verbose=F, preProcess = c("center", "scale"))
rffit
stopCluster(cl)
```

Examine quality of prediction
```{r}
predrf<-predict(rffit, newdata=test)
confusionMatrix(predrf, test$classe)
```

Apply machine learning algorithm to the 20 test cases available in the test data
```{r}
pml_testingTidy["problem_id"] <- NULL
test_cases<-predict(rffit, newdata=pml_testingTidy)
test_cases
```